# Aim: Now I am adding the reworked_annotations_alternative to the vis_mod object. 
At least here I won't have mixed annotations in the same layer but between layers. I am gonna process each annotation layer separately. This way I can isolate the specific annotation in a layer and correspoding spots (eg. epithelial invasive) to do pairwise comparisons. 

```{r}
# Load packages
library(semla)
library(ggplot2)
library(tidyverse)
library(dplyr)
```

Load vis_mod object 
```{r}
# Load data (Seurat object with corrected image size)
vis_mod <- readRDS(file = paste0("/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/output/rds_objects/", "all_samples_cropped_img.rds"))
```

```{r}
#Load annotations
load_annotations <- function(layer_name) {
  # List directories where annotation files are stored
  dirs <- list.dirs("/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/hard_vs_soft_coverslip/results/path_annotations/reworked_annotations_alternative",
                    recursive = FALSE, full.names = TRUE)
  
  # Process annotations for each sample section
  anno.df <- map_dfr(unique(vis_mod$sample_id), function(id) {
    # Find annotation files matching the current sample and layer
    fs <- list.files(str_subset(dirs, pattern = paste0(id, "$")), 
                     pattern = paste0(layer_name, "_reworked.csv"),  # Match specific layer
                     full.names = TRUE, recursive = TRUE)
    
    # Read annotation file (assumes one file per sample per layer)
    if (length(fs) == 0) {
      return(NULL)  # Skip if no file found
    }
    
    anno_df <- read_csv(fs)
    
    # Ensure consistent column names
    colnames(anno_df) <- c("barcode", "annotation")  # Adjust if needed
    
    # Adjust barcode suffix to match vis_mod
    suf <- which(unique(vis_mod$sample_id) == id)
    anno_df <- anno_df %>%
      mutate(barcode = str_replace(barcode, "\\d+$", as.character(suf)))
    
    return(anno_df)
  })
  
  return(anno.df)
}

```

```{r}
# Load each annotation layer separately
epithelial_annotations <- load_annotations("epithelial")
immune_annotations <- load_annotations("immune")
stroma_annotations <- load_annotations("stroma")
```

```{r}
# Add one layer at a time to the vis_mod Seurat object 
# Add epithelial annotations
vis_mod <- AddMetaData(vis_mod, epithelial_annotations %>% column_to_rownames("barcode"), col.name = "Epithelial_Anno")

# Add immune annotations
vis_mod <- AddMetaData(vis_mod, immune_annotations %>% column_to_rownames("barcode"), col.name = "Immune_Anno")

# Add stroma annotations
vis_mod <- AddMetaData(vis_mod, stroma_annotations %>% column_to_rownames("barcode"), col.name = "Stroma_Anno")
```


```{r}
#Because the desmoplastic stroma annotation intermigles with the epithelial layer I am gonna rename as NA
vis_mod@meta.data <- vis_mod@meta.data %>%
  mutate(Stroma_Anno = ifelse(Stroma_Anno == "desmoplastic", NA, Stroma_Anno))
```

```{r}
#save Seurat object with loaded annotations
saveRDS(vis_mod,
        file = paste0("/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/output/rds_objects/", "vis_mod_anno.rds"))
```


Load vis_mod object 
```{r}
# Load data (new Seurat object modified by Sophie) with my added annotations
vis_mod <- readRDS(file = paste0("/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/output/rds_objects/", "vis_mod_anno.rds"))
```

1.QC

```{r}
#This coerces the Spatial assay into a classic Seurat RNA assay, and stores it as "RNA3" in your object.
names(vis_mod@assays)
vis_mod[["RNA3"]] <- as(object = vis_mod[["Spatial"]], Class = "Assay")

names(vis_mod@assays) # check that an additional assay called "RNA3" is created
```

```{r}
# Mitochondrial
mt.genes <- grep(pattern = "^MT-", x = rownames(vis_mod), value = TRUE)
vis_mod$percent.mito <- (Matrix::colSums(vis_mod@assays$RNA@counts[mt.genes,])/Matrix::colSums(vis_mod@assays$RNA@counts))*100
head(vis_mod$percent.mito) # There is only 1 to 5% of mitochrondial genes cause these are not included in the 10x gen pannel.

# Ribosomal
# Collect all genes coding for ribosomal proteins
rp.genes <- grep(pattern = "^RP[SL]", x = rownames(vis_mod), value = TRUE)
vis_mod$percent.ribo <- (Matrix::colSums(vis_mod@assays$RNA@counts[rp.genes,])/Matrix::colSums(vis_mod@assays$RNA@counts))*100
head(vis_mod$percent.ribo)  # There is no ribosomal genes cause these are not included in the 10x gen pannel.

# Percentage hemoglobin genes - includes all genes starting with HB except HBP.
#Collect hemoglobin protein coding genes
hemo.genes <- grep(pattern = "^HB[^(P|E|S)]", x = rownames(vis_mod), value = TRUE)
vis_mod$percent.hemo <- (Matrix::colSums(vis_mod@assays$RNA@counts[hemo.genes,])/Matrix::colSums(vis_mod@assays$RNA@counts))*100
head(vis_mod$percent.hemo) # The percentage of hemo genes is like 0%

# Percentage for some platelet markers
plat.genes <- grep(pattern = "PECAM1|PF4", x = rownames(vis_mod), value = TRUE)
vis_mod$percent.plat <- (Matrix::colSums(vis_mod@assays$RNA@counts[plat.genes,])/Matrix::colSums(vis_mod@assays$RNA@counts))*100
head(vis_mod$percent.plat) #The percentage is almost 0
```

Plot QC

```{r, fig.width = 15, fig.height = 8}
feats <- c("nFeature_Spatial", "nCount_Spatial", "percent.mito", "percent.hemo", "percent.plat")
QC_all_samples <- VlnPlot(vis_mod, group.by = "sample_id", split.by = "method", features = feats, pt.size = 0.1, ncol = 5)
```

Plot spatially to understand ViolinPlots

```{r}
m <- MapFeatures(vis_mod_norm, features = "percent.mito", cols = c("lightgray", "mistyrose", "red", "dark red", "black"), pt.size = 0.1, ncol = 4) #something is happening with the pt.size
m
```


Yay! Save as PNG 
```{r}
png_res <- 1200
png(filename = "/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/results_LA/QC_all_samples.png", width = 15*png_res, height = 8*png_res, res = png_res);print(QC_all_samples);dev.off()
```

#### Save object

```{r}
saveRDS(vis_mod,
        file = paste0("/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/output/rds_objects/", "vis_mod_anno_QC.rds"))
```

2. Normalize Data and Find Variable Genes. 
```{r}
# Normalize data and find top variable features
vis_mod_norm <- vis_mod |> 
  NormalizeData() |> 
#We need to define which features/genes are important in our dataset
  FindVariableFeatures()
```
Scaling the data by centering the data and scaling (Z-score)

```{r}
#Since each gene has a different expression level, it means that genes with higher expression values will naturally have higher variation that will be captured by PCA. This means that we need to somehow give each gene a similar weight when performing PCA
vis_mod_norm <- ScaleData(vis_mod_norm)
```

3. Dimensionality reduction Run PCA() on variable genes to capture main sources of variation

```{r}
vis_mod_norm <- RunPCA(vis_mod_norm, npcs = 50, verbose = F)
```

```{r}
ElbowPlot(vis_mod_norm) #which help you determine how many PCs actually capture meaningful biological variation (often 10–30).
```
Run UMAP

```{r}
vis_mod_norm <- RunUMAP(
  vis_mod_norm,
  reduction = "pca",
  dims = 1:10, #if elbow plot suggested 10 PCs!
  n.components = 2,
  n.neighbors = 50,
  min.dist = 0.2,
  verbose = TRUE
)
# see ?RunUMAP for more info
```
Visualize UMAP

```{r}
UMAP <- DimPlot(
  vis_mod_norm,
  reduction = "umap",
  group.by = "sample_id"
) + ggplot2::ggtitle("UMAP on PCA")
```
Yay! Save as PNG 
```{r}
png_res <- 1200
png(filename = "/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/results_LA/UMAP.png", width = 15*png_res, height = 8*png_res, res = png_res);print(UMAP);dev.off()
```

Testing if normalizing and scaling helps at all. It did!!! Same scale :D
```{r}
KRT5_plot <- MapFeatures(vis_mod_norm,
            colors = RColorBrewer::brewer.pal(n = 9, name = "Reds"),
            feature = "KRT5", 
            image_use = "raw",
            ncol = 4,
            pt_alpha = 0.5,
            scale_alpha = FALSE, 
            max_cutoff = 0.95,
            add_scalebar = T, scalebar_position = c(0.8,0.9),
            pt_stroke = NA,
            pt_size = 1,
            slot = "counts")
```

Yay! Save as PNG 
```{r}
png_res <- 1200
png(filename = "/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/results_LA/KRT5_plot_on_scaled_data.png", width = 15*png_res, height = 8*png_res, res = png_res);print(KRT5_plot);dev.off()
```

#### Save object vis_mod_anno_QC_normalized 

```{r}
saveRDS(vis_mod_norm,
        file = paste0("/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/output/rds_objects/new_object/", "vis_mod_norm.rds"))
```

Trying mapping annotations on tissue 
```{r}
annotation_colors <- c('#ffe119', '#46f0f0',"#eeeeee")
# Here we use the & operator from the patchwork R package to add a theme
# You can find more details in the 'advanced' tutorial
MapLabels(vis_mod, column_name = "Stroma_Anno", colors = annotation_colors, pt_size = 1.3, pt_alpha = 0.5, section_number = 8, ncol = 1) &
  theme(legend.position = "right")

p <- MapLabels(vis_mod, column_name = "Immune_Anno", pt_size = 1.3, pt_alpha = 0.5, section_number = 8, ncol = 1) &
  theme(legend.position = "right")
print(p)
```

```{r}
# Here we use the & operator from the patchwork R package to add a theme
# You can find more details in the 'advanced' tutorial
MapLabels(vis_mod, column_name = "Stroma_Anno", pt_size = 1.3, pt_alpha = 0.5, colors = annotation_colors, section_number = 7, ncol = 1) &
  theme(legend.position = "right")
```

```{r}
annotation_colors <- c('#ffe119', '#46f0f0',"#eeeeee")
# Here we use the & operator from the patchwork R package to add a theme
# You can find more details in the 'advanced' tutorial
MapLabels(vis_mod, column_name = "Stroma_Anno", pt_size = 1.3, pt_alpha = 0.5, colors = annotation_colors, section_number = 1, ncol = 1) &
  theme(legend.position = "right")
```

```{r}
annotation_colors <- c('#ffe119', '#46f0f0',"#eeeeee")
# Here we use the & operator from the patchwork R package to add a theme
# You can find more details in the 'advanced' tutorial
MapLabels(vis_mod, column_name = "Stroma_Anno", pt_size = 1.3, pt_alpha = 0.5, colors = annotation_colors, section_number = 2, ncol = 1) &
  theme(legend.position = "right")
```

```{r}
annotation_colors <- c('#ffe119', '#46f0f0',"#eeeeee")
# Here we use the & operator from the patchwork R package to add a theme
# You can find more details in the 'advanced' tutorial
MapLabels(vis_mod, column_name = "Stroma_Anno", colors = annotation_colors, pt_size = 2, pt_alpha = 0.5, section_number = 3, ncol = 1) &
  theme(legend.position = "right")
```

```{r}
annotation_colors <- c('#ffe119', '#46f0f0',"#eeeeee")
# Here we use the & operator from the patchwork R package to add a theme
# You can find more details in the 'advanced' tutorial
MapLabels(vis_mod, column_name = "Stroma_Anno", pt_size = 2, pt_alpha = 0.5, colors = annotation_colors, section_number = 4, ncol = 1) &
  theme(legend.position = "right")
```
```{r}
annotation_colors <- c('#ffe119', '#46f0f0',"#eeeeee")
# Here we use the & operator from the patchwork R package to add a theme
# You can find more details in the 'advanced' tutorial
MapLabels(vis_mod, column_name = "Stroma_Anno", pt_size = 2, pt_alpha = 0.5, colors = annotation_colors, section_number = 5, ncol = 1) &
  theme(legend.position = "right")
```

```{r}
annotation_colors <- c('#ffe119', '#46f0f0',"#eeeeee")
# Here we use the & operator from the patchwork R package to add a theme
# You can find more details in the 'advanced' tutorial
MapLabels(vis_mod, column_name = "Stroma_Anno", pt_size = 2, pt_alpha = 0.5, colors = annotation_colors, section_number = 6, ncol = 1) &
  theme(legend.position = "right")
```
Let's see how the epithelial and immune annotation looks like 

```{r}
# Here we use the & operator from the patchwork R package to add a theme
# You can find more details in the 'advanced' tutorial
MapLabels(vis_mod, column_name = "Epithelial_Anno", ncol = 4) &
  theme(legend.position = "right")
```

```{r}
# Here we use the & operator from the patchwork R package to add a theme
# You can find more details in the 'advanced' tutorial
MapLabels(vis_mod, column_name = "Immune_Anno", ncol = 4) &
  theme(legend.position = "right")
```

```{r}
# Define a named vector linking section_id to method
method_map <- c("4335_3D_R1L_s1" = "hard_cover", 
                "4335_3D_R1L_s2" = "soft_cover",
                "4393_2H_R1S" = "soft_cover", 
                "4393_2H_R1S_rep" = "hard_cover", 
                "4419_3F_R1S" = "soft_cover", 
                "4419_3F_R1S_rep" = "hard_cover",
                "4515_4L_R1L" = "soft_cover",
                "4515_4L_R1L_rep" = "hard_cover")
                
# Assign method dynamically based on section_id in vis_mod
vis_mod@meta.data$method <- method_map[vis_mod@meta.data$sample_id]

# Check if order is consistent
table(vis_mod@meta.data$method)
table(vis_mod@meta.data$sample_id, vis_mod@meta.data$method)

# Use the following code to reorder the section_id


vis_mod@meta.data$sample_id <- factor(vis_mod@meta.data$sample_id, 
                                       levels = c("4335_3D_R1L_s1", 
                                                  "4335_3D_R1L_s2", 
                                                  "4393_2H_R1S_rep", 
                                                  "4393_2H_R1S", 
                                                  "4419_3F_R1S_rep", 
                                                  "4419_3F_R1S", 
                                                  "4515_4L_R1L_rep", 
                                                  "4515_4L_R1L"))
# Check if order is consistent
table(vis_mod@meta.data$sample_id, vis_mod@meta.data$method)
```

#### Save vis_mod object with annotations and method

```{r}
saveRDS(vis_mod,
        file = paste0("/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/output/rds_objects/", "vis_mod_anno.rds"))
```


Violin plots per layer 
Epithelial 

```{r}
vis_mod_ano <- readRDS(paste0("/Users/leirealonso/Documents/ST/Project_Garvan_institute/Human_Breast_Cancer_Atlas/hard_vs_soft_cover/for_manuscripts/analysis/output/rds_objects/", "vis_mod_ano.rds"))
```

```{r}
# Subset based on the annotation using an expession
vis_mod_epithelial <- SubsetSTData(vis_mod, expression = Epithelial_Anno %in% c("invasive", "in_situ", "normal"))
```

```{r}
# Now, plot.df1 will contain the metadata with epithelial anno, method, and sample_id columns as factors, alongside the newly created log_nCount column. You can then use this plot.df1 data frame for plotting or further analysis.
plot.df1 <- vis_mod_epithelial@meta.data %>% 
  mutate(
    log_nCount = log1p(nCount_Spatial),
    Epithelial_Anno = as.factor(Epithelial_Anno),
    method = as.factor(method),
    sample_id = as.factor(sample_id)
  )
#str(plot.df1): Displays the structure of plot.df1, showing each column's name, data type (e.g., factor, numeric, etc.), and a preview of its values.
```

```{r}
p <- ggplot(vis_mod_epithelial@meta.data, aes(y = nCount_Spatial, x = sample_id, fill = method)) +
  geom_violin(trim = FALSE, scale = "width") +
  scale_y_log10() +
  labs(title = "Log-transformed UMI Count Distribution in Epithelial Layer",
       x = "Sample ID",
       y = "Log-transformed UMI Count") +
  geom_boxplot(width = 0.1, outlier.shape = NA) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(p)

```


```{r}
# Filter the data to include only one block and using grouped annotations
  plot.df.4515 <- plot.df1.group %>% 
  filter(block_id %in% c("4515_4L_R1L")) # Focus on block 4515
# Plot the data using the filtered grouped annotations for all sections
ggplot(plot.df.4515, aes(x = grouped_annotation, y = nCount_Spatial, fill = method)) +
  geom_violin(trim = FALSE, scale = "width", width = 0.7) +  
  facet_wrap(~ block_id, ncol = 2) +  # Facet by block_id
  scale_y_log10() +  # Log-transform y-axis
  theme_minimal() +  # Clean up the plot appearance
  labs(title = "Comparison of UMI Counts by Annotation and Method",
       y = "UMI Counts", x = "Annotation") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),  # Increase title size
    axis.title = element_text(size = 14),  # Increase axis title size
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),  # Rotate x-axis labels and increase size
    axis.text.y = element_text(size = 12),  # Increase y-axis text size
    legend.position = "right",  # Position legend on the right
    strip.text = element_text(size = 14)  # Increase facet label size
  ) 
```




